---
title: "Summary of the data from clinicaltrials.gov"
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: 
    toc: true
    toc_depth: 2
    reference_docx: rmarkdown-styles-reference.docx
---

```{r setup, include=FALSE}
# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
library(klaR) # for k-modes clustering
library(dplyr)
library(broom)
library(stringr)
library(visdat) # for missing data
library(ggplot2)
library(gridExtra)
library(ggupset) # to plot combination of study characteristics
library(pander)
library(summarytools)
# global options for summary tools
st_options(plain.ascii = FALSE,       # Always use this option in Rmd documents
            style = "rmarkdown",        # This too
            round.digits = 0, 
            headings = FALSE, 
            footnote = NA,             # Avoids footnotes which would clutter the results
            subtitle.emphasis = FALSE  # Improves layout with some rmardown themes
) 

# graphics things:
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7")

# get the data
load('data/clinicaltrials_analysis_ready.RData') # from 1_process_clintrials_data.R 
```

There were `r length(excluded)` studies excluded.
The total number of remaining studies is `r format(nrow(studies),big.mark=',')`. 
The data were extracted from clinicaltrials.gov on `r censor.date`.

# Missing data

The columns are variables and the rows are observations.
The grey areas show missing data.
The plot shows a random sample of 10,000 observations.

```{r, fig.width=8}
for.missing = dplyr::select(studies, -id, -starts_with('n_')) %>%
  sample_n(size=10000) # too big so take a random sample
vis_dat(for.missing)
```

# Registration and dates

## Registration dates

```{r, results='asis'}
dates = select(studies, posted, updated, completed, results) %>%
  tidyr::gather(key='date', value='res') %>%
  group_by(date) %>%
  filter(!is.na(res)) %>% # remove missing
  summarise(n=n(), 
            median = median(res), 
            q1 = as.Date(quantile(unclass(res), 0.25), origin = "1970-01-01"),
            q3 = as.Date(quantile(unclass(res), 0.75), origin = "1970-01-01")) %>%
  mutate(date = factor(date, levels=c('posted','updated','completed','results')),
         n = format(n,big.mark = ','),
         median = as.Date(median, origin='1970-01-01')) %>% # For row ordering
  arrange(date)
pander(dates, style='simple')
```

The summary statistics are the median and inter-quartile range.

# Inclusions/Exclusions

## Gender included

```{r, results='asis'}
with(studies, freq(gender, cumul = FALSE, report.nas =TRUE))
```

## Age included

#### Minimum age

```{r, results='asis'}
with(studies, freq(age_min_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Minimum age limit in years

```{r}
# split by age
to.plot = filter(studies, !is.na(age_min)) 
# round to years for over 1
over_1 = filter(to.plot, age_min >=1) %>%
  mutate(age_min = round(age_min)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_min < 1) %>%
  mutate(age_min = floor(age_min*12)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Minimum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  xlab('Minimum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for studies that had a minimum age limit.

#### Maximum age

```{r, results='asis'}
with(studies, freq(age_max_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Maximum age limit in years

```{r}
# split by age
to.plot = filter(studies, !is.na(age_max)) 
# round to years for over 1
over_1 = filter(to.plot, age_max >=1) %>%
  mutate(age_max = round(age_max)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_max < 1) %>%
  mutate(age_max = floor(age_max*12)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Maximum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  xlab('Maximum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for studies that had a maximum age limit.

# Study characteristics

## Study type

```{r, results='asis'}
with(studies, freq(study_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Purpose

```{r, results='asis'}
with(studies, freq(purpose, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Masking

```{r, results='asis'}
with(studies, freq(masking, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Assignment

```{r, results='asis'}
with(studies, freq(assignment, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Phase

```{r, results='asis'}
with(studies, freq(phase, cumul = FALSE, report.nas =TRUE))
```

## Number of outcomes

```{r, results='asis'}
# could add 10% and 90%?
to.summary = select(studies, n_primary, n_secondary)
descr(to.summary, stats=c('fivenum'))
```

Number of primary and secondary outcomes.
A small number of studies had a huge number of outcomes.


# Study status

```{r, results='asis'}
with(studies, freq(status, cumul = FALSE, report.nas =FALSE, order='freq')) 
```


# Length of description

```{r, results='asis'}
to.summary = select(studies, description_nchar)
descr(to.summary, stats=c('N.valid','Min','Q1','Med','Q3','Max'))
```

This is the number of characters written in the description section.

# Funding

### Lead sponsor type

```{r, results='asis'}
with(studies, freq(lead_sponsor_class, cumul = FALSE, report.nas =FALSE, order='freq'))
```


# Sample size

## Actual and achieved sample size

```{r, results='asis'}
sample = select(studies, status, sample_size, sample_size_type) %>%
  filter(!status=='Withdrawn' | sample_size_type =='Anticipated') # exclude those withdrawn with an actual sample size (often zero)
stby(data = sample, 
           INDICES = sample$sample_size_type, 
           FUN = descr, stats = "fivenum", transpose = TRUE)
zero.studies = nrow(filter(studies, sample_size == 0))
```

The table shows summary statistics.

There were `r zero.studies` studies with a zero sample size.


## Histogram of sample size

```{r, fig.width=8}
exclude_large = filter(studies, sample_size < 2000) # exclude extremely large genetics study

#
hplot = ggplot(exclude_large, aes(x=sample_size))+
  geom_histogram(fill='chocolate1')+
  xlab('Sample size')+
  g.theme+
  facet_wrap(~sample_size_type, scales='free_y')
hplot
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size under 2,000.


### Histogram looking for digit preference in target sample size

```{r, fig.width=8}
# exclude huge studies
exclude_large = filter(studies, sample_size <= 200) # exclude extremely large genetics study
# frequency
digits = group_by(exclude_large, sample_size_type, sample_size) %>%
  summarise(n=n()) %>%
  ungroup() 
top_five = group_by(digits, sample_size_type) %>% # per type (facet)
  arrange(-n) %>%
  slice(1:5) %>% # top 5 per facet
  ungroup()
# plot
hplot = ggplot(digits, aes(x=sample_size, y=n))+
  geom_bar(stat="identity", col='chocolate2')+
  xlab('Sample size')+
  ylab('Count')+
  geom_label(data=top_ten, aes(x=sample_size, y=n, label=sample_size))+
  g.theme+
  facet_wrap(~sample_size_type, scales='free_y')
hplot
jpeg('figures/digit_preference_clintrials.jpg', width=5, height=4, units='in', res=300)
print(hplot)
invisible(dev.off())
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size of 200 or fewer.

There is a strong digit preference at multiples of ten. The modal sample size is `r digits$samplesize_target[1]`. The top five sample sizes are shown for each sample size type.


## Multiple regression model of sample size

```{r}
# check if results are already available
is_results = length(dir('results', pattern = 'clintrials_sample_size')) > 0
if(is_results == TRUE){
  load('results/clintrials_sample_size.RData')
}
if(is_results == FALSE){
  source('2_elasticnet_model_samplesize_clintrial.R')
}
```

```{r, fig.width=8, fig.height=9}
# make table of estimates and names, and add reference groups
# some results put in order of high to low
table_names = read.table(sep='!', header=TRUE, stringsAsFactors = FALSE, text='
xaxis!group!term!label!reference
1!continuous!date!Trend per 5 years!FALSE
2!continuous!n_primary!Double the number of primary outcomes!FALSE
3!continuous!n_secondary!Double the number of secondary outcomes!FALSE
4!continuous!n_condition!Double the Number of conditions!FALSE
5!continuous!n_arms!Double the number of arms!FALSE
6!Gender!genderMale!Males!FALSE
7!Gender!genderFemale!Females!FALSE
8!Gender!genderAll!All!TRUE
9!Age limit!age_limitOver 18!Over 18!FALSE
10!Age limit!age_limitExactly 18!Exactly 18!FALSE
11!Age limit!age_limitUnder 18!Under 18!FALSE
12!Age limit!age_limitNo limit!No limit!TRUE
13!Type!study_typeInterventional!Interventional!TRUE
14!Type!study_typeObservational!Observational!FALSE
15!Phase!phaseEarly Phase 1!Early Phase 1!FALSE
16!Phase!phasePhase 1!Phase 1!FALSE
17!Phase!phasePhase 1, Phase 2!Phase 1/2!FALSE
18!Phase!phasePhase 2!Phase 2!FALSE
19!Phase!phasePhase 2, Phase 3!Phase 2/3!FALSE
20!Phase!phasePhase 3!Phase 3!TRUE
21!Phase!phasePhase 4!Phase 4!FALSE
22!Phase!phaseNot Applicable!Not Applicable!FALSE
23!Sponsor!lead_sponsor_classOTHER!Other!TRUE
24!Sponsor!lead_sponsor_classFED!FED!FALSE
25!Sponsor!lead_sponsor_classINDUSTRY!Industry!FALSE
26!Sponsor!lead_sponsor_classNETWORK!Network!FALSE
27!Sponsor!lead_sponsor_classNIH!NIH!FALSE
28!Purpose!purposeDevice Feasibility!Device Feasibility!FALSE
29!Purpose!purposeBasic Science!Basic Science!FALSE
30!Purpose!purposeTreatment!Treatment!TRUE
31!Purpose!purposeOther!Other!FALSE
32!Purpose!purposeDiagnostic!Diagnostic!FALSE
33!Purpose!purposePrevention!Prevention!FALSE
34!Purpose!purposeHealth Services Research!Health Services Research!FALSE
35!Purpose!purposeScreening!Screening!FALSE
36!Masking!maskingTriple!Triple!FALSE
37!Masking!maskingTriple!Double!TRUE
38!Masking!maskingNone (Open Label)!Open label!FALSE
39!Assignment!assignmentCrossover!Crossover!FALSE
40!Assignment!assignmentSingle Group!Single group!FALSE
41!Assignment!assignmentSequential!Sequential!FALSE
42!Assignment!assignmentParallel!Parallel!TRUE
43!Assignment!assignmentFactorial!Factorial!FALSE
44!Allocation!allocationNon-Randomized!Non-Randomized!FALSE
45!Allocation!allocationNon-Randomized!Randomized!TRUE
') %>%
  arrange(xaxis) # sort in order for labels
# make reference groups
reference_groups = filter(table_names, reference==TRUE)  %>%
 mutate(estimate=0)

# add above to estimates
add_ests = filter(table_names, reference==FALSE) %>%
  left_join(standard_model_ests, by='term') %>%
  bind_rows(reference_groups) %>% # add references
  mutate(
    # back transform to relative change:
    estimate = exp(estimate), 
    conf.low = exp(conf.low),
    conf.high = exp(conf.high)
  ) %>%
  dplyr::select(xaxis, estimate, conf.low, conf.high, reference) %>%
  mutate(reference = as.numeric(reference=='TRUE') + 1) # to mark reference point
axis_labels = table_names$label

# quick check that all estimates are in the reference table (should only be intercept); and vice versa
check_merge = anti_join(standard_model_ests, table_names, by='term')
check_merge_reverse = anti_join(table_names, standard_model_ests, by='term') %>%
  filter(reference == FALSE)

# labels for groups
group_labels = group_by(table_names, group) %>%
  summarise(n=n(), meanx=mean(xaxis), maxx=max(xaxis)) %>%
  ungroup() %>%
  filter(n > 1) %>%
  mutate(
    estimate = max(add_ests$conf.high, na.rm=T), conf.low=0, conf.high=0, reference=1,
    group = ifelse(group=='continuous', 'Continuous\noutcomes', group)) 
dotted.lines = group_labels$maxx + 0.5 # dotted lines to split groups
# text for axis labels
text1 = data.frame(xaxis=0, estimate=1, conf.low=0, conf.high=0, reference=1, label='Increase')
text2 = data.frame(xaxis=0, estimate=1, conf.low=0, conf.high=0, reference=1, label='Decrease')
# plot
star.wars.relative = ggplot(data=add_ests, aes(x=xaxis, y=estimate, ymin=conf.low, ymax=conf.high, shape=factor(reference)))+
  geom_point(size=2)+
  scale_shape_manual(NULL, values=c(19,17))+ # triangle and circle
  geom_errorbar(width=0, size=1.02)+
  geom_hline(lty=2, yintercept=1)+
  geom_vline(lty=3, xintercept=c(5.5, dotted.lines))+ # breaks between categorical variables
  geom_text(data=text1, aes(x=xaxis, y=estimate, label =label), adj=-0.05, col=grey(0.5), size=3.5)+
  geom_text(data=text2, aes(x=xaxis, y=estimate, label =label), adj=1.05, col=grey(0.5), size=3.5)+
  geom_text(data=group_labels, aes(x=meanx, y=estimate, label =group), adj=1, col=grey(0.5))+
  scale_x_continuous(expand=c(0.01,0.01), breaks=1:length(axis_labels), labels=axis_labels, limits=c(0, length(axis_labels)))+
  scale_y_continuous(breaks=seq(0,4,1), minor_breaks = seq(0,4,0.5), limits=c(0.21, NA))+ # avoid showing 0 on RR axis
  ylab('Relative change in sample size')+
  xlab('')+
  theme_bw()+
  theme(legend.position = 'none',
        text=element_text(size=13), 
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())+
  coord_flip() # to 
star.wars.relative
# export to tall-thin plot
jpeg('figures/clintrials_sample_size_relative.jpg', width=6, height=7.5, units='in', res=300)
print(star.wars.relative)
invisible(dev.off())
```


## Common characteristics

Below we show the top ten characteristics of studies in terms of study type, purpose and masking.

```{r, fig.width=8, fig.height=6}
# get combinations 
combs <-  dplyr::select(studies, purpose, masking, study_type, id)  %>%
  as_tibble() %>%
  tidyr::gather(key='characteristic', value='response', -id) %>%
  filter(!is.na(response),
         response != 'Not Applicable') %>%
  dplyr::select(-characteristic)
# now make list from group responses
list_resp = group_by(combs, id) %>%
  summarise(responses = list(response))
# plot
cplot = ggplot(list_resp, aes(x = responses)) +
    geom_bar(aes(y=..count../sum(..count..)), fill = "indianred3") +
    theme_bw() +
    xlab("Study characteristics") +
    ylab("Proportion of studies") +
    scale_x_upset(n_intersections = 10)+
    g.theme+
  theme(text=element_text(size=14),
        plot.margin = margin(t = 0, r = 0, b = 0, l = 60, unit = "pt")) # extend margin for labels
cplot
```


## Clustering similar studies

We use a k-modes clustering algorithm to look for studies with similar characteristics.
We assumed there are 10 clusters.

```{r cluster}
set.seed(780734)
for.cluster = select(studies, lead_sponsor_class, study_type, purpose, allocation, phase, assignment, masking, gender, age_max_type, age_min_type ) %>% 
  mutate_all(tidyr::replace_na, replace = 'Missing')  # replace all missing
cl = kmodes(for.cluster, 10)
# check which variables do not differ and remove them
diffs = sapply(sapply(cl$modes, unique), length)
removed = names(cl$modes)[diffs == 1]
to_table = cl$modes[, diffs > 1 ] %>%
  mutate(count=cl$size) %>%
  arrange(-count) %>%
  mutate(count = format(count, big.mark = ','))
pander(to_table, style='simple')
```
