---
title: "ANZCTR summary"
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: 
    toc: true
    toc_depth: 2
    reference_docx: rmarkdown-styles-reference.docx
---

```{r setup, include=FALSE}
# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
library(dplyr)
library(broom)
library(stringr)
library(visdat) # for missing data
library(ggplot2)
library(gridExtra)
library(ggupset) # to plot combination of study characteristics
library(flextable)
library(summarytools)
# global options for summary tools
st_options(plain.ascii = FALSE,       # Always use this option in Rmd documents
            style = "rmarkdown",        # This too
            round.digits = 0, 
            headings = FALSE, 
            footnote = NA,             # Avoids footnotes which would clutter the results
            subtitle.emphasis = FALSE  # Improves layout with some rmardown themes
) 
# for PLOS
library(extrafont)
loadfonts(device = "win")

# graphics things:
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7")

# get the data
load('data/AnalysisReady.RData') # from 0_read_data_anzctr.R 
source('99_functions.R')
```

These summary statistics are for trials registered in ANZCTR and not clinicaltrials.gov.

The total number of studies is `r format(nrow(studies),big.mark=',')`. 
The data were extracted from ANZCTR on `r censor.date`.

# Missing data

The columns are variables and the rows are observations.
The grey areas show missing data.

```{r, fig.width=8}
for.missing = dplyr::select(studies, -pub, -time, -ID, -id, -starts_with('n_'))
vis_dat(for.missing)
```

# Registration and dates

## Registration dates

```{r, results='asis'}
dates = select(studies, submitted, approved, update) %>%
  tidyr::gather(key='date', value='res') %>%
  group_by(date) %>%
  filter(!is.na(res)) %>% # remove missing
  summarise(n=n(), 
            median = median(res), 
            q1 = as.Date(quantile(unclass(res), 0.25), origin = "1970-01-01"),
            q3 = as.Date(quantile(unclass(res), 0.75), origin = "1970-01-01")) %>%
  mutate(date = factor(date, levels=c('submitted','approved','update')),
         n = format(n,big.mark = ','),
         median = as.Date(median, origin='1970-01-01'),
         median = format(median, '%b %Y'), # just month and year
         q1 = format(q1, '%b %Y'),
         q3 = format(q3, '%b %Y')) %>% 
  arrange(date)
ftab = flextable(dates) %>%
  theme_box() %>%
  autofit()
ftab
```

The summary statistics are the median and inter-quartile range.

## Anticipated start date

```{r}
stats = filter(studies, !is.na(start_anticipated)) %>%
  mutate(start_anticipated = as.numeric(start_anticipated)) %>%
        summarise(min=min(start_anticipated),
                  q1=quantile(start_anticipated, probs=0.25),
                  median=median(start_anticipated),
                  q3=quantile(start_anticipated, probs=0.75),
                  max=max(start_anticipated)) %>%
  mutate_all(funs(as.Date(., origin='1970-01-01'))) %>%
  mutate_all(funs(format(., '%b %Y'))) # change all to month - year
ftab = flextable(stats) %>%
  theme_box() %>%
  autofit()
ftab
```

The summary statistics are the minimum, inter-quartile range, median and maximum.

# Inclusions/Exclusions

## Gender included

```{r, results='asis'}
with(studies, freq(gender, cumul = FALSE, report.nas =TRUE))
```

## Age included

#### Minimum age

```{r, results='asis'}
with(studies, freq(age_min_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Minimum age limit in years

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_min)) 
# round to years for over 1
over_1 = filter(to.plot, age_min >=1) %>%
  mutate(age_min = round(age_min)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_min < 1) %>%
  mutate(age_min = floor(age_min*12)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Minimum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  xlab('Minimum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a minimum age limit.
By far the most common minimum age is 18.
The oldest exclusion for minimum age was `r max(to.plot$age_min)` years.

#### Maximum age

```{r, results='asis'}
with(studies, freq(age_max_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Maximum age limit in years

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_max)) 
# round to years for over 1
over_1 = filter(to.plot, age_max >=1) %>%
  mutate(age_max = round(age_max)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_max < 1) %>%
  mutate(age_max = floor(age_max*12)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Maximum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  xlab('Maximum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```


This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a maximum age limit.
The oldest exclusion for maximum age was `r max(to.plot$age_max)` years.

There is a strong digit preference in the age limit with spikes in multiples of five years.

# Study characteristics

## Study type

```{r, results='asis'}
with(studies, freq(study_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Study type over time

```{r}
to_plot = mutate(studies,
                 year = as.numeric(format(submitted, '%Y'))) %>% # make year
  group_by(year, study_type) %>%
  tally()
gplot = ggplot(data=to_plot, aes(x=year, y=n, fill=study_type))+
  geom_bar(stat='identity')+
  scale_fill_manual(NULL, values=c('darkseagreen3','darkgoldenrod2'))+
  theme_bw()+
  ylab('Number of studies')+
  xlab('Year submitted')+
  theme(legend.position = c(0.15,0.8))
gplot
```

## Purpose

```{r, results='asis'}
with(studies, freq(purpose, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Masking

```{r, results='asis'}
with(studies, freq(masking, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Intervention code

```{r, results='asis'}
with(studies, freq(intervention_code, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Control

```{r, results='asis'}
with(studies, freq(control, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Phase

```{r, results='asis'}
with(studies, freq(phase, cumul = FALSE, report.nas =TRUE))
```

## Endpoint

```{r, results='asis'}
with(studies, freq(endpoint, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Number of outcomes

```{r, results='asis'}
to.summary = select(studies, n_primary, n_secondary)
descr(to.summary, stats=c('fivenum'))
```

Number of primary and secondary outcomes.

# Health conditions

## Condition code #1

```{r, results='asis'}
with(studies, freq(ccode1, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

## Condition code #2 (top 10)

```{r, results='asis'}
freq = group_by(studies, ccode2) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(studies,
                  ccode2 = ifelse(ccode2 %in% freq$ccode2, ccode2, 'Other')) # replace conditions outside top 10 with 'other'
with(to.table, freq(ccode2, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

There are `r length(unique(studies$ccode2))` condition codes, hence we just show the top ten.
Conditions outside the top ten have been grouped together as 'other'.

## Condition as free text (top 10)

```{r, results='asis'}
freq = group_by(studies, condition) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(studies,
                  condition = ifelse(condition %in% freq$condition, condition, 'Other')) # replace conditions outside top 10 with 'other'
with(to.table, freq(condition, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

There are `r format(length(unique(studies$condition)), big.mark=',')` conditions as free text, hence we just show the top ten.

This field is free text and so relies on what researchers write. Some categories could likely be combined, e.g., "prostate cancer" and "prostate cancer patients".

# Study status

## Study status

```{r, results='asis'}
with(studies, freq(study_status, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

## Provisional

```{r, results='asis'}
with(studies, freq(provisional, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

## Study-related publication listed

```{r, results='asis'}
with(studies, freq(pub, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

This is just a check of whether there is a study publication listed. It does not check whether it is the main publication of the study results.

# Sponsors and funding

## Top institutes

```{r, results='asis'}
freq = group_by(studies, address) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(studies,
                  address = ifelse(address %in% freq$address, address, 'Other')) # replace those outside top 10 with 'other'
with(to.table, freq(address, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

There are `r format(length(unique(studies$address)), big.mark=',')` institutes, hence we just show the top ten.

## Funding

### Number of funders

```{r, results='asis'}
with(studies, freq(n_funding, cumul = FALSE, report.nas =FALSE))
```

### Funding type

```{r, results='asis'}
with(funding_data, freq(type, cumul = FALSE, report.nas =FALSE, order='freq'))
```

This includes multiple results per study.

### Funders (top 10)

```{r, results='asis'}
freq = group_by(funding_data, name) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(funding_data,
                  name = ifelse(name %in% freq$name, name, 'Other')) # replace funders outside top 10 with 'other'
with(to.table, freq(name, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

This includes multiple results per study.

There are `r format(length(unique(funding_data$name)), big.mark=',')` funders, hence we just show the top ten.


# Sample size

## Target and achieved sample size

```{r, results='asis'}
sample = select(studies, samplesize_target, samplesize_actual) %>%
  rename('Target' = 'samplesize_target',
         'Achieved' = 'samplesize_actual')
descr(sample, stats='fivenum')
zero.studies = nrow(filter(studies, samplesize_target == 0))
```

The table shows summary statistics.

There were `r zero.studies` studies with a zero target sample size.


## Histogram of sample size

```{r}
exclude_large = filter(studies, samplesize_target < 2000) # exclude extremely large genetics study
hplot = ggplot(exclude_large, aes(x=samplesize_target))+
  geom_histogram(fill='chocolate1')+
  xlab('Target sample size')+
  g.theme
hplot
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size under 2,000.

### Histogram looking for digit preference in target sample size

```{r, fig.width=8}
#
exclude_large = filter(studies, samplesize_target <= 200) # exclude extremely large genetics study
# 
digits = group_by(exclude_large, samplesize_target) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # top 10
# plot
hplot = ggplot(exclude_large, aes(x=samplesize_target))+
  geom_histogram(fill='chocolate2', breaks=0:200)+
  xlab('Target sample size')+
  geom_label(data=digits, aes(x=samplesize_target, y=n, label=samplesize_target))+
  g.theme
hplot
jpeg('figures/digit_preference.jpg', width=5, height=4, units='in', res=300)
print(hplot)
invisible(dev.off())
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size of 200 or fewer.

There is a strong digit preference at multiples of ten. The modal sample size is `r digits$samplesize_target[1]`.


## Multiple regression model of target and actual sample size

```{r, include=FALSE}
## load data/results
# check if results are already available
is_results = length(dir('results', pattern = 'ANZCTR_sample_size')) > 0
if(is_results == TRUE){
  load('results/ANZCTR_sample_size.RData')
}
if(is_results == FALSE){
  source('2_elasticnet_model_samplesize_ANZCTR.R')
}

# get the labels
load('data/labels.RData') # from 0_labels.R
```

We used elastic net. 
Selected parsimonious model as minimum cross-validated error plus 1 standard error.
Checked the final model for colinearity using the variance-inflation factor (VIF).
There were variables that were colinear with study type (Observational/Interventional) and this was part of the reason for running these models separately.

We log-transformed (base e) the sample size because of a very strong positive skew.
The plot shows the estimated mean and 95% confidence interval on the untransformed scale.

### Plot of percent differences for the actual sample size

```{r, fig.width=8, fig.height=9.5}
to_plot = plot_function(indata=all_ests, which_outcome='actual', table_names=table_names_anzctr)
to_plot
# export to tall-thin plot
jpeg('figures/ANZCTR_sample_size_relative_actual.jpg', width=6.5, height=9.5, units='in', res=600)
print(to_plot)
invisible(dev.off())
# for PLOS
tiff('figures/Fig2.tif', width=6.5, height=8.75, units='in', res=600, family='Times New Roman', compression = 'lzw') # 
print(to_plot)
invisible(dev.off())
```

The number of studies was `r format(filter(numbers, outcome=='actual', study_type=='Observational')  %>% pull(n_model), big.mark=',')` for the observational data and `r format(filter(numbers, outcome=='actual', study_type=='Interventional')  %>% pull(n_model), big.mark=',')` for the interventional data.

### Plot of relative differences for the target sample size

```{r, fig.width=8, fig.height=9.5}
to_plot = plot_function(indata=all_ests, which_outcome='target', table_names = table_names_anzctr)
to_plot
# export to tall-thin plot
jpeg('figures/ANZCTR_sample_size_relative_target.jpg', width=6.5, height=9.5, units='in', res=600)
print(to_plot)
invisible(dev.off())
# for PLOS
tiff('figures/Fig3.tif', width=6.5, height=8.75, units='in', res=600, family='Times New Roman', compression = 'lzw') # 
print(to_plot)
invisible(dev.off())
```

The number of studies was `r format(filter(numbers, outcome=='target', study_type=='Observational')  %>% pull(n_model), big.mark=',')` for the observational data and `r format(filter(numbers, outcome=='target', study_type=='Interventional')  %>% pull(n_model), big.mark=',')` for the interventional data.



### Check residuals from regression model

#### a) Target sample size

```{r, fig.width=7, fig.height=3.5}
otype = 'target'
all_res = NULL
for (model in c('Interventional','Observational')){ # get residuals for both types
  model_comb = paste(otype, model, sep=' ')
  res = data.frame(res = resid(standard_models[[model_comb]]))
  frame = data.frame(model=model, res=res)
  all_res = bind_rows(all_res, frame)
}
rplot = ggplot(all_res, aes(x=res))+
  geom_histogram(fill='firebrick3')+
  g.theme+
  xlab('Residual (log-scale)')+
  ylab('Count')+
  facet_wrap(~model, scales='free_y')
rplot
# r-squared
# rsq = cor(log2(for.model$samplesize_target), fitted(smodel)) ^2
```

The scales on the y-axes differ across panels.

#### b) Actual sample size

```{r, fig.width=7, fig.height=3.5}
otype = 'actual'
all_res = NULL
for (model in c('Interventional','Observational')){ # get residuals for both types
  model_comb = paste(otype, model, sep=' ')
  res = data.frame(res = resid(standard_models[[model_comb]]))
  frame = data.frame(model=model, res=res)
  all_res = bind_rows(all_res, frame)
}
rplot = ggplot(all_res, aes(x=res))+
  geom_histogram(fill='firebrick3')+
  g.theme+
  xlab('Residual (log-scale)')+
  ylab('Count')+
  facet_wrap(~model, scales='free_y')
rplot
```


