---
title: "Summary of the data and models from clinicaltrials.gov"
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: 
    toc: true
    toc_depth: 2
    reference_docx: rmarkdown-styles-reference.docx
---

```{r setup, include=FALSE}
# using formatting in Word document (see above)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
library(dplyr)
library(broom)
library(stringr)
library(visdat) # for missing data
library(ggplot2)
library(gridExtra)
library(ggupset) # to plot combination of study characteristics
library(flextable)
library(summarytools)
# global options for summary tools
st_options(plain.ascii = FALSE,       # Always use this option in Rmd documents
            style = "rmarkdown",        # This too
            round.digits = 0, 
            headings = FALSE, 
            footnote = NA,             # Avoids footnotes which would clutter the results
            subtitle.emphasis = FALSE  # Improves layout with some rmardown themes
) 
source('99_functions.R') # for plot_function
# for PLOS
library(extrafont)
loadfonts(device = "win")

# graphics things:
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7")

# get the data
load('data/clinicaltrials_analysis_ready.RData') # from 1_process_clintrials_data.R 
```

There were `r length(excluded)` studies excluded.
The total number of remaining studies is `r format(nrow(studies),big.mark=',')`. 
The data were extracted from clinicaltrials.gov on `r censor.date`.

# Missing data

The columns are variables and the rows are observations.
The grey areas show missing data.
The plot shows a random sample of 10,000 observations.

```{r, fig.width=8}
for.missing = dplyr::select(studies, -id, -starts_with('n_')) %>%
  sample_n(size=10000) # too big so take a random sample
vis_dat(for.missing)
```

# Registration and dates

## Registration dates

```{r, results='asis'}
dates = dplyr::select(studies, posted, updated) %>%
  tidyr::gather(key='date', value='res') %>%
  group_by(date) %>%
  filter(!is.na(res)) %>% # remove missing
  summarise(n=n(), 
            median = median(res), 
            q1 = as.Date(quantile(unclass(res), 0.25), origin = "1970-01-01"),
            q3 = as.Date(quantile(unclass(res), 0.75), origin = "1970-01-01")) %>%
  mutate(date = factor(date, levels=c('posted','updated','completed','results')),
         n = format(n,big.mark = ','),
         median = as.Date(median, origin='1970-01-01'), # For row ordering
         median = format(median, '%b %Y'), # just month and year
         q1 = format(q1, '%b %Y'),
         q3 = format(q3, '%b %Y')) %>% 
  arrange(date)
ftab = flextable(dates) %>%
  theme_box() %>%
  autofit()
ftab
```

The summary statistics are the median and inter-quartile range.

# Inclusions/Exclusions

## Gender included

```{r, results='asis'}
with(studies, freq(gender, cumul = FALSE, report.nas =TRUE))
```

## Age included

#### Minimum age

```{r, results='asis'}
with(studies, freq(age_min_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Minimum age limit in years

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_min)) 
# round to years for over 1
over_1 = filter(to.plot, age_min >=1) %>%
  mutate(age_min = round(age_min)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_min < 1) %>%
  mutate(age_min = floor(age_min*12)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Minimum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  xlab('Minimum age (years)')+
  ylab('Count')+
  g.theme+
  theme(plot.margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt"))+ # extend right margin
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a minimum age limit.
By far the most common minimum age is 18.
The oldest exclusion for minimum age was `r max(to.plot$age_min)` years.

#### Maximum age

```{r, results='asis'}
with(studies, freq(age_max_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Maximum age limit in years

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_max)) 
# round to years for over 1
over_1 = filter(to.plot, age_max >=1) %>%
  mutate(age_max = round(age_max)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_max < 1) %>%
  mutate(age_max = floor(age_max*12)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Maximum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  xlab('Maximum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a maximum age limit.
The oldest exclusion for maximum age was `r max(to.plot$age_max)` years.

# Study characteristics

## Study type

```{r, results='asis'}
with(studies, freq(study_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Purpose

```{r, results='asis'}
with(studies, freq(purpose, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Masking

```{r, results='asis'}
with(studies, freq(masking, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Assignment

```{r, results='asis'}
with(studies, freq(assignment, cumul = FALSE, report.nas =TRUE, order='freq'))
```

## Phase

```{r, results='asis'}
with(studies, freq(phase, cumul = FALSE, report.nas =TRUE))
```

## Number of outcomes

```{r, results='asis'}
# could add 10% and 90%?
to.summary = select(studies, n_primary, n_secondary)
descr(to.summary, stats=c('fivenum'))
```

Number of primary and secondary outcomes.
A small number of studies had a huge number of outcomes.

# Study status

```{r, results='asis'}
with(studies, freq(status, cumul = FALSE, report.nas =FALSE, order='freq')) 
```


# Funding

### Lead sponsor type

```{r, results='asis'}
with(studies, freq(lead_sponsor_class, cumul = FALSE, report.nas =FALSE, order='freq'))
```


# Sample size

## Actual and achieved sample size

```{r, results='asis'}
sample = select(studies, status, sample_size, sample_size_type) %>%
  filter(!status=='Withdrawn' | sample_size_type =='Anticipated') # exclude those withdrawn with an actual sample size (often zero)
stby(data = sample, 
           INDICES = sample$sample_size_type, 
           FUN = descr, stats = "fivenum", transpose = TRUE)
zero.studies = nrow(filter(studies, sample_size == 0))
```

The table shows summary statistics.

There were `r zero.studies` studies with a zero sample size.


## Histogram of sample size

```{r, fig.width=8}
exclude_large = filter(studies, sample_size < 2000) # exclude extremely large genetics study

#
hplot = ggplot(exclude_large, aes(x=sample_size))+
  geom_histogram(fill='chocolate1')+
  xlab('Sample size')+
  g.theme+
  facet_wrap(~sample_size_type, scales='free_y')
hplot
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size under 2,000.


### Histogram looking for digit preference in target sample size

```{r, fig.width=8}
# exclude huge studies
exclude_large = filter(studies, sample_size <= 200) # exclude extremely large genetics study
# frequency
digits = group_by(exclude_large, sample_size_type, sample_size) %>%
  summarise(n=n()) %>%
  ungroup() 
top_five = group_by(digits, sample_size_type) %>% # per type (facet)
  arrange(-n) %>%
  slice(1:5) %>% # top 5 per facet
  ungroup()
# plot
hplot = ggplot(digits, aes(x=sample_size, y=n))+
  geom_bar(stat="identity", col='chocolate2')+
  xlab('Sample size')+
  ylab('Number of studies')+
  geom_label(data=top_five, aes(x=sample_size, y=n, label=sample_size))+
  g.theme+
  facet_wrap(~sample_size_type, scales='free_y')
hplot
jpeg('figures/digit_preference_clintrials.jpg', width=5, height=4, units='in', res=300)
print(hplot)
invisible(dev.off())
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size of 200 or fewer.

There is a strong digit preference at multiples of ten. The modal sample size is `r top_five$sample_size[1]`. The top five sample sizes are shown for each sample size type.


## Multiple regression model of sample size

```{r}
# load the multiple regression results
# check if results are already available
is_results = length(dir('results', pattern = 'clintrials_sample_size')) > 0
if(is_results == TRUE){
  load('results/clintrials_sample_size.RData')
}
if(is_results == FALSE){
  source('2_elasticnet_model_samplesize_clintrial.R')
}
# get the labels
load('data/labels.RData') # from 0_labels.R
```

### a) target sample size

```{r, fig.width=8, fig.height=9}
to_plot = plot_function(indata = all_ests, 
                        which_outcome = 'target',  
                        ljust=0.35, # move legend at top
                        table_names = table_names_clintrials, 
                        x_limits = seq(-100,300,100),
                        minor_breaks = seq(-100,350,50))
to_plot
# export to tall-thin plot
jpeg('figures/clintrials_sample_size_relative_target.jpg', width=6.5, height=9.5, units='in', res=600)
print(to_plot)
invisible(dev.off())
## export to PLOS ready plot
tiff('figures/Fig4.tif', width=6.5, height=8.75, units='in', res=600, family='Times New Roman', compression = 'lzw') # 
print(to_plot)
invisible(dev.off())
```

### b) actual sample size

```{r, fig.width=8, fig.height=9}
to_plot = plot_function(indata = all_ests, 
                        which_outcome = 'actual', 
                        ljust=0.35,
                        table_names = table_names_clintrials, 
                        x_limits = seq(-100,300,50),
                        minor_breaks = seq(-100,350,50))
to_plot
# export to tall-thin plot
jpeg('figures/clintrials_sample_size_relative_actual.jpg', width=6.5, height=9.8, units='in', res=600)
print(to_plot)
invisible(dev.off())
## export to PLOS ready plot
tiff('figures/Fig5.tif', width=6.5, height=8.75, units='in', res=600, family='Times New Roman', compression = 'lzw') # 
print(to_plot)
invisible(dev.off())
```


### Check residuals from regression models

```{r, fig.width=7, fig.height=3.5}
rsq = all_res = NULL
for (outcome_type in c('target','actual')){ # get residuals for both 
  for (model in c('Interventional','Observational')){ # get residuals for both 
    mtype = paste(outcome_type, model)
    mtype_nice = paste(stringr::str_to_sentence(outcome_type), model, sep=', ')
    res = data.frame(res = resid(standard_models[[mtype]]))
    frame = data.frame(model=mtype_nice, res=res)
    all_res = bind_rows(all_res, frame)
    # r-squared
    #r = cor(log2(for.model$samplesize_target), fitted(standard_models[[model]])) ^2
    #rsq = c(rsq, r)
  }
}
rplot = ggplot(all_res, aes(x=res))+
  geom_histogram(fill='firebrick3')+
  g.theme+
  xlab('Residual (log-scale)')+
  ylab('Count')+
  facet_wrap(~model, scales='free')
rplot
```
