---
title: "Supplement 1: Descriptive statistics and descriptions of missing data for the two databases"
output:
  word_document: 
    toc: true
    toc_depth: 2
    reference_docx: rmarkdown-styles-reference.docx
---

```{r setup, include=FALSE}
# 
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation
library(dplyr)
library(stringr)
library(visdat) # for missing data
library(naniar) # for missing data
library(ggplot2)
library(gridExtra)
library(ggupset) # to plot combination of study characteristics
library(flextable)
library(summarytools)
# global options for summary tools
st_options(plain.ascii = FALSE,       # Always use this option in Rmd documents
            style = "rmarkdown",        # This too
            round.digits = 0, 
            headings = FALSE, 
            footnote = NA,             # Avoids footnotes which would clutter the results
            subtitle.emphasis = FALSE  # Improves layout with some rmardown themes
) 

# graphics things:
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#999999", "#CC79A7")

#
load('data/labels.RData') # from 0_labels.R
```

<!--- page break --->
###### page break

# clinicaltrials.gov

```{r, include=FALSE}
# get the data and process ready for the regression model
load('data/clinicaltrials_analysis_plus.RData') # from 2a_update_clintrials_data.R 
source('3_prep_clintrials.R')
```

There were `r format(nrow(excluded),big.mark=',')` studies excluded. 
The total number of remaining studies is `r format(nrow(studies),big.mark=',')`. 
The data were extracted from clinicaltrials.gov on `r format(censor.date, '%d-%b-%Y')`.

## Missing data (clinicaltrials.gov)

The plots below summarise the missing data.
The columns are variables and the rows are observations.
The grey areas show missing data.
The coloured plots show a random sample of 10,000 studies as the data are too large to show every study.

```{r, include=FALSE}
# function needed to re-create missing
back.to.missing = function(x){
  x[x=='Missing'] = NA
  return(x)
}
#
for.missing = dplyr::select(for.model, -id, -date, -platform_trial) %>% # date is generated; platform trial not used
  mutate_if(is.factor, back.to.missing) # convert "Missing" back to NA
```

```{r, fig.width=8}
ten_thousand = sample_n(for.missing, size=10000) # too big so take a random sample
gplot = vis_dat(ten_thousand)
gplot + ylab('Studies')
```

### Counts of missing data (clinicaltrials.gov)

The plots below shows the counts of missing data using all studies.

```{r}
cplot = gg_miss_var(for.missing, show_pct = FALSE)
# get slightly nicer variable names
levels = cplot$data$variable
labels = str_replace_all(levels, '_', ' ')
cplot + xlab('')+
  scale_x_discrete(breaks=levels, labels=labels)
```

The total number of studies in `r format(nrow(for.missing), big.mark=',')`.

Overall the amount of missing data was small. The most frequent missing variable was the study purpose which is the main objective of the intervention(s) being evaluated by the clinical trial. This was missing for `r round(100* sum(is.na(for.missing$purpose))/nrow(for.missing))`\% of studies.

The logical variables cannot be missing by design, for example 'longitudinal' was created by whether the word "longitudinal" was mentioned in the detailed study description or not. 

Many variables were mandatory and so cannot be missing, e.g., study status.

## Registration and dates (clinicaltrials.gov)

### Registration dates (clinicaltrials.gov)

```{r, results='asis'}
dates = dplyr::select(studies, posted, updated) %>%
  tidyr::gather(key='date', value='res') %>%
  group_by(date) %>%
  filter(!is.na(res)) %>% # remove missing
  summarise(n=n(), 
            median = median(res), 
            q1 = as.Date(quantile(unclass(res), 0.25), origin = "1970-01-01"),
            q3 = as.Date(quantile(unclass(res), 0.75), origin = "1970-01-01")) %>%
  mutate(date = factor(date, levels=c('posted','updated','completed','results')),
         n = format(n,big.mark = ','),
         median = as.Date(median, origin='1970-01-01'), # For row ordering
         median = format(median, '%b %Y'), # just month and year
         q1 = format(q1, '%b %Y'),
         q3 = format(q3, '%b %Y')) %>% 
  arrange(date)
ftab = flextable(dates) %>%
  theme_box() %>%
  autofit()
ftab
```

The summary statistics are the median and inter-quartile range.

## Inclusions/Exclusions (clinicaltrials.gov)

### Gender included (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(gender, cumul = FALSE, report.nas =TRUE))
```

### Age included (clinicaltrials.gov)

##### Minimum age (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(age_min_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```


##### Minimum age limit in years (clinicaltrials.gov)

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_min)) 
# round to years for over 1
over_1 = filter(to.plot, age_min >=1) %>%
  mutate(age_min = round(age_min)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_min < 1) %>%
  mutate(age_min = floor(age_min*12)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Minimum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  xlab('Minimum age (years)')+
  ylab('Count')+
  g.theme+
  theme(plot.margin = margin(t = 0, r = 10, b = 0, l = 0, unit = "pt"))+ # extend right margin
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```


This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a minimum age limit.
By far the most common minimum age is 18.
The oldest exclusion for minimum age was `r max(to.plot$age_min)` years.

##### Maximum age (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(age_max_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

##### Maximum age limit in years (clinicaltrials.gov)

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_max)) 
# round to years for over 1
over_1 = filter(to.plot, age_max >=1) %>%
  mutate(age_max = round(age_max)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_max < 1) %>%
  mutate(age_max = floor(age_max*12)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Maximum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  xlab('Maximum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a maximum age limit.
The oldest exclusion for maximum age was `r max(to.plot$age_max)` years.

### Healthy volunteers included (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(volunteers, cumul = FALSE, report.nas =TRUE))
```

## Study characteristics (clinicaltrials.gov)

### Purpose (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(purpose, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Masking (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(masking, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Assignment (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(assignment, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Allocation (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(allocation, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Phase (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(phase, cumul = FALSE, report.nas =TRUE))
```

### Adaptive/platform trial (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(adaptive_trial, cumul = FALSE, report.nas =FALSE))
```

Was an adaptive or platform trial mentioned in the study description.

### Longitudinal (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(longitudinal, cumul = FALSE, report.nas =FALSE))
```

Was longitudinal mentioned in the study description.

### Number of outcomes (clinicaltrials.gov)

```{r, results='asis'}
# could add 10% and 90%?
to.summary = select(studies, n_primary, n_secondary)
descr(to.summary, stats=c('fivenum'))
```

Number of primary and secondary outcomes.
A small number of studies had a huge number of outcomes.

## Study status (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(status, cumul = FALSE, report.nas =FALSE, order='freq')) 
```


## Funding (clinicaltrials.gov)

### Lead sponsor type (clinicaltrials.gov)

```{r, results='asis'}
with(studies, freq(lead_sponsor_class, cumul = FALSE, report.nas =FALSE, order='freq'))
```

## Sample size (clinicaltrials.gov)

### Target and achieved sample size (clinicaltrials.gov)

```{r, results='asis'}
sample = select(studies, samplesize_target, samplesize_actual) %>%
  rename('Target' = 'samplesize_target',
         'Achieved' = 'samplesize_actual')
descr(sample, stats='fivenum')
zero.studies = nrow(filter(studies, samplesize_target == 0))
```

The table shows summary statistics.

There were `r zero.studies` studies with a zero target sample size.

### Histogram of sample size (clinicaltrials.gov)

```{r}
exclude_large = filter(studies, samplesize_target < 2000) # exclude extremely large genetics study
hplot = ggplot(exclude_large, aes(x=samplesize_target))+
  geom_histogram(fill='chocolate1')+
  xlab('Target sample size')+
  g.theme
hplot
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size under 2,000.

### Histogram looking for digit preference in target sample size (clinicaltrials.gov)

```{r, fig.width=8}
# exclude huge studies
exclude_large = filter(studies, samplesize_target <= 200) # exclude extremely large genetics study
# frequency
digits = group_by(exclude_large, samplesize_target) %>% # per type (facet)
  tally() 
top_five = arrange(digits, -n) %>%
  slice(1:5) %>% # top 5 per facet
  ungroup()
# plot
hplot = ggplot(digits, aes(x=samplesize_target, y=n))+
  geom_bar(stat="identity", col='chocolate2')+
  xlab('Sample size')+
  ylab('Number of studies')+
  geom_label(data=top_five, aes(x=samplesize_target, y=n, label=samplesize_target))+
  g.theme
hplot
jpeg('figures/digit_preference_clintrials.jpg', width=5, height=4, units='in', res=300)
print(hplot)
invisible(dev.off())
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size of 200 or fewer.

There is a strong digit preference at multiples of ten. The modal sample size is `r top_five$sample_size[1]`. The top five sample sizes are shown for each sample size type.

<!--- page break --->
###### 

# ANZCTR

```{r, include=FALSE}
# get the data
load('data/AnalysisReady.RData') # from 0_read_data_anzctr.R 
# prep the data as per the models
source('3_prep_ANZCTR.R')
```

These summary statistics are for trials registered in ANZCTR and not clinicaltrials.gov.

The total number of studies is `r format(nrow(studies),big.mark=',')`. 
The data were extracted from ANZCTR on `r format(censor.date, '%d-%b-%Y')`.

## Missing data (ANZCTR)

The columns are variables and the rows are observations.
The grey areas show missing data.

```{r}
for.missing = dplyr::select(for.model, -id, -ID, -pub, -time, -date, -address, -start_anticipated, -end_date, -n_ethics_committees, -number, -ethics_review, -age_min, -age_max) %>% # date is generated; drop variables not used in this analysis
  mutate_if(is.factor, back.to.missing) # convert "Missing" back to NA
```

```{r, fig.width=8}
gplot = vis_dat(for.missing)
gplot + ylab('Studies')
```


### Counts of missing data (ANZCTR)

The plots below shows the counts of missing data.

```{r}
cplot = gg_miss_var(for.missing, show_pct = FALSE)
# get slightly nicer variable names
levels = cplot$data$variable
labels = str_replace_all(levels, '_', ' ')
cplot + xlab('')+
  scale_x_discrete(breaks=levels, labels=labels)
```


## Registration and dates (ANZCTR)

### Registration dates (ANZCTR)

```{r, results='asis'}
dates = select(studies, submitted, approved, update) %>%
  tidyr::gather(key='date', value='res') %>%
  group_by(date) %>%
  filter(!is.na(res)) %>% # remove missing
  summarise(n=n(), 
            median = median(res), 
            q1 = as.Date(quantile(unclass(res), 0.25), origin = "1970-01-01"),
            q3 = as.Date(quantile(unclass(res), 0.75), origin = "1970-01-01")) %>%
  mutate(date = factor(date, levels=c('submitted','approved','update')),
         n = format(n,big.mark = ','),
         median = as.Date(median, origin='1970-01-01'),
         median = format(median, '%b %Y'), # just month and year
         q1 = format(q1, '%b %Y'),
         q3 = format(q3, '%b %Y')) %>% 
  arrange(date)
ftab = flextable(dates) %>%
  theme_box() %>%
  autofit()
ftab
```

The summary statistics are the median and inter-quartile range.

### Anticipated start date (ANZCTR)

```{r}
stats = filter(studies, !is.na(start_anticipated)) %>%
  mutate(start_anticipated = as.numeric(start_anticipated)) %>%
        summarise(min=min(start_anticipated),
                  q1=quantile(start_anticipated, probs=0.25),
                  median=median(start_anticipated),
                  q3=quantile(start_anticipated, probs=0.75),
                  max=max(start_anticipated)) %>%
  mutate_all(funs(as.Date(., origin='1970-01-01'))) %>%
  mutate_all(funs(format(., '%b %Y'))) # change all to month - year
ftab = flextable(stats) %>%
  theme_box() %>%
  autofit()
ftab
```

The summary statistics are the minimum, inter-quartile range, median and maximum.

## Inclusions/Exclusions (ANZCTR)

### Gender included (ANZCTR)

```{r, results='asis'}
with(studies, freq(gender, cumul = FALSE, report.nas =TRUE))
```

### Age included (ANZCTR)

#### Minimum age (ANZCTR)

```{r, results='asis'}
with(studies, freq(age_min_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Minimum age limit in years (ANZCTR)

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_min)) 
# round to years for over 1
over_1 = filter(to.plot, age_min >=1) %>%
  mutate(age_min = round(age_min)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_min < 1) %>%
  mutate(age_min = floor(age_min*12)) %>%
  group_by(age_min) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Minimum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_min, y=n))+
  geom_histogram(fill='darkseagreen', stat='identity')+
  xlab('Minimum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a minimum age limit.
By far the most common minimum age is 18.
The oldest exclusion for minimum age was `r max(to.plot$age_min)` years.

#### Maximum age (ANZCTR)

```{r, results='asis'}
with(studies, freq(age_max_type, cumul = FALSE, report.nas =TRUE, order='freq'))
```

#### Maximum age limit in years (ANZCTR)

```{r, fig.width=7}
# split by age
to.plot = filter(studies, !is.na(age_max)) 
# round to years for over 1
over_1 = filter(to.plot, age_max >=1) %>%
  mutate(age_max = round(age_max)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# round to months for under 1
under_1 = filter(to.plot, age_max < 1) %>%
  mutate(age_max = floor(age_max*12)) %>%
  group_by(age_max) %>%
  summarise(n = n()) %>%
  ungroup()
# separate plots
hplot1 = ggplot(under_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  scale_x_continuous(breaks=0:12)+
  xlab('Maximum age (months)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age under 1')
hplot2 = ggplot(over_1, aes(x=age_max, y=n))+
  geom_histogram(fill='dodgerblue', stat='identity')+
  xlab('Maximum age (years)')+
  ylab('Count')+
  g.theme+
  ggtitle('Age 1 and over')
# combine plot
grid.arrange(hplot1, hplot2, ncol=2)
```

This plot is just for the `r format(nrow(to.plot), big.mark=',')` studies that had a maximum age limit.
The oldest exclusion for maximum age was `r max(to.plot$age_max)` years.

There is a strong digit preference in the age limit with spikes in multiples of five years.

### Healthy volunteers included (ANZCTR)

```{r, results='asis'}
with(studies, freq(volunteers, cumul = FALSE, report.nas =TRUE))
```

## Study characteristics (ANZCTR)

### Purpose (ANZCTR)

```{r, results='asis'}
with(studies, freq(purpose, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Assignment (ANZCTR)

```{r, results='asis'}
with(studies, freq(assignment, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Allocation (ANZCTR) 

```{r, results='asis'}
with(studies, freq(allocation, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Masking (ANZCTR) 

```{r, results='asis'}
with(studies, freq(masking, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Intervention code (ANZCTR)

```{r, results='asis'}
with(studies, freq(intervention_code, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Endpoint (ANZCTR)

```{r, results='asis'}
with(studies, freq(endpoint, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Phase (ANZCTR) 

```{r, results='asis'}
with(studies, freq(phase, cumul = FALSE, report.nas =TRUE))
```

### Control (ANZCTR) 

```{r, results='asis'}
with(studies, freq(control, cumul = FALSE, report.nas =TRUE, order='freq'))
```

### Number of outcomes (ANZCTR)

```{r, results='asis'}
to.summary = select(studies, n_primary, n_secondary)
descr(to.summary, stats=c('fivenum'))
```

Number of primary and secondary outcomes.

## Health conditions (ANZCTR)

### Condition code #1 (ANZCTR)

```{r, results='asis'}
with(studies, freq(ccode1, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

### Condition code #2, top 10 (ANZCTR)

```{r, results='asis'}
freq = group_by(studies, ccode2) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(studies,
                  ccode2 = ifelse(ccode2 %in% freq$ccode2, ccode2, 'Other')) # replace conditions outside top 10 with 'other'
with(to.table, freq(ccode2, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

There are `r length(unique(studies$ccode2))` condition codes, hence we just show the top ten.
Conditions outside the top ten have been grouped together as 'other'.

### Condition as free text, top 10 (ANZCTR)

```{r, results='asis'}
freq = group_by(studies, condition) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(studies,
                  condition = ifelse(condition %in% freq$condition, condition, 'Other')) # replace conditions outside top 10 with 'other'
with(to.table, freq(condition, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

There are `r format(length(unique(studies$condition)), big.mark=',')` conditions as free text, hence we just show the top ten.

This field is free text and so relies on what researchers write. Some categories could likely be combined, e.g., "prostate cancer" and "prostate cancer patients".

## Study status (ANZCTR)

### Study status (ANZCTR)

```{r, results='asis'}
with(studies, freq(study_status, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

## Sponsors and funding (ANZCTR)

### Top institutes (ANZCTR)

```{r, results='asis'}
freq = group_by(studies, address) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(studies,
                  address = ifelse(address %in% freq$address, address, 'Other')) # replace those outside top 10 with 'other'
with(to.table, freq(address, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

There are `r format(length(unique(studies$address)), big.mark=',')` institutes, hence we just show the top ten.

### Funding (ANZCTR)

#### Number of funders (ANZCTR)

```{r, results='asis'}
with(studies, freq(n_funding, cumul = FALSE, report.nas =FALSE))
```

#### Funding type (ANZCTR)

```{r, results='asis'}
with(funding_data, freq(type, cumul = FALSE, report.nas =FALSE, order='freq'))
```

This includes multiple results per study.

#### Funders, top 10 (ANZCTR)

```{r, results='asis'}
freq = group_by(funding_data, name) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # just top ten
to.table = mutate(funding_data,
                  name = ifelse(name %in% freq$name, name, 'Other')) # replace funders outside top 10 with 'other'
with(to.table, freq(name, cumul = FALSE, report.nas =FALSE, order='freq')) 
```

This includes multiple results per study.

There are `r format(length(unique(funding_data$name)), big.mark=',')` funders, hence we just show the top ten.


## Sample size (ANZCTR)

### Target and achieved sample size (ANZCTR)

```{r, results='asis'}
sample = select(studies, samplesize_target, samplesize_actual) %>%
  rename('Target' = 'samplesize_target',
         'Achieved' = 'samplesize_actual')
descr(sample, stats='fivenum')
zero.studies = nrow(filter(studies, samplesize_target == 0))
```

The table shows summary statistics.

There were `r zero.studies` studies with a zero target sample size.


### Histogram of sample size (ANZCTR)

```{r}
exclude_large = filter(studies, samplesize_target < 2000) # exclude extremely large genetics study
hplot = ggplot(exclude_large, aes(x=samplesize_target))+
  geom_histogram(fill='chocolate1')+
  xlab('Target sample size')+
  g.theme
hplot
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size under 2,000.

#### Histogram looking for digit preference in target sample size (ANZCTR)

```{r, fig.width=8}
#
exclude_large = filter(studies, samplesize_target <= 200) # exclude extremely large genetics study
# 
digits = group_by(exclude_large, samplesize_target) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  arrange(-n) %>%
  slice(1:10) # top 10
# plot
hplot = ggplot(exclude_large, aes(x=samplesize_target))+
  geom_histogram(fill='chocolate2', breaks=0:200)+
  xlab('Target sample size')+
  geom_label(data=digits, aes(x=samplesize_target, y=n, label=samplesize_target))+
  g.theme
hplot
jpeg('figures/digit_preference.jpg', width=5, height=4, units='in', res=300)
print(hplot)
invisible(dev.off())
```

The plot above is for the `r format(nrow(exclude_large), big.mark=',')` studies with a target sample size of 200 or fewer.

There is a strong digit preference at multiples of ten. The modal sample size is `r digits$samplesize_target[1]` and this is dominated by withdrawn studies.
