---
title: "Tables of predictor variables in the two trial registries"
output: pdf_document
classoption: landscape
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation

library(dplyr)
library(stringr)
library(flextable)
## get the data
# a) clintrials.gov
load('results/var_list_clintrials.RData') # from 2_elasticnet_model_samplesize_clintrial.R
# create observational / interventional variable
int = filter(vars, stype=='Interventional') %>% select(-stype, -database)
obs = filter(vars, stype=='Observational') %>% select(-stype, -database)
vars = full_join(int, obs, by=c('variables','categories','log')) %>%
  mutate(stype = 'Interventional',
         stype = ifelse(variables %in% int$variables & variables %in% obs$variables, 'Interventional/Observational', stype),
         stype = ifelse(!variables %in% int$variables & variables %in% obs$variables, 'Observational', stype))
clintrials_vars = select(vars, variables, stype, categories, log) # column ordering

# b) ANZCTR
load('results/var_list_anzctr.RData') # from 2_elasticnet_model_samplesize_ANZCTR.R
# create observational / interventional variable
int = filter(vars, stype=='Interventional') %>% select(-stype, -database)
obs = filter(vars, stype=='Observational') %>% select(-stype, -database)
vars = full_join(int, obs, by=c('variables','categories','log')) %>%
  mutate(stype = 'Interventional',
         stype = ifelse(variables %in% int$variables & variables %in% obs$variables, 'Interventional/Observational', stype),
         stype = ifelse(!variables %in% int$variables & variables %in% obs$variables, 'Observational', stype))
anzctr_vars = select(vars, variables, stype, categories, log) # column ordering
```


```{r, include=FALSE}
## add nicer variable names
load('data/labels.RData') # from '0_labels.R'
# add a few more labels
more_labels = read.table(sep=',', header=TRUE, stringsAsFactors = FALSE, text='
term,label
age_limit_max,Maximum age limit
age_limit_min,Minimum age limit
volunteers,Healthy volunteers
provisional,Provisional results
ccode1,Area
study_status,Study status
intervention_code,Intervention
study_design_time,Study design
study_design_observational,Study time
lead_sponsor_class,Lead sponsor
')
table_names_clintrials = bind_rows(table_names_clintrials, more_labels)
table_names_anzctr = bind_rows(table_names_anzctr, more_labels)

# merge with data
clintrials_vars = left_join(clintrials_vars, table_names_clintrials, by=c("variables" = "term")) %>%
  mutate(Variable = ifelse(is.na(label)==TRUE, str_to_sentence(variables), label)) %>% # sentence case if no label
  select(Variable, stype, categories, log) %>%
  rename('Study types' = 'stype')
anzctr_vars = left_join(anzctr_vars, table_names_anzctr, by=c("variables" = "term")) %>%
  mutate(Variable = ifelse(is.na(label), str_to_sentence(variables), label)) %>%
  select(Variable, stype, categories, log) %>%
  rename('Study types' = 'stype')

```

\thispagestyle{empty} 

The tables below show the predictor variables available for the two trial registries that were used in the multiple regression models. 

The "study types" column shows if the variable was available for the Interventional and/or Observational models. 

The "categories" column gives the categories for factors. 

The "log" column indicates whether the variable was log-transformed (base 2). 

## clintrials.gov

\thispagestyle{empty} 

```{r}
# add whether they are mandatory or optional?
ftab = flextable(clintrials_vars) %>%
  fontsize(part="body", size=10) %>%
  width(width=c(2.9, 2, 4.6, 0.3)) %>%
  theme_box()
ftab
```

## ANZCTR

\thispagestyle{empty} 


```{r}
ftab = flextable(anzctr_vars) %>%
#  fontsize(size = 8, part = "all") %>%
  width(width=c(2, 2, 6, 0.4)) %>%
  theme_box()
ftab
```

Study status was only used for the models of actual sample size and not the target sample size, this is because the study status was likely to change after the target sample size was set.
