---
title: "Variables excluded from final models by the elastic net selection"
output: pdf_document
classoption: portrait
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
options(width=1000) # Wide pages
options(scipen=999) # avoid scientific presentation

library(dplyr)
library(stringr)
library(flextable)

# get the labels
load('data/labels.RData')
```


# clintrials.gov

```{r, include=FALSE}
# get the results
load('results/clintrials_sample_size.RData') # from 2_elasticnet_model_samplesize_clintrial.R
```

## Target sample size for observational studies

```{r}
vars_out = filter(variables_excluded,
              studytype == 'Observational',
              outcome == 'target') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

## Target sample size for interventional studies

```{r}
vars_out = filter(variables_excluded,
              studytype == 'Interventional',
              outcome == 'target') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

## Actual sample size for observational studies

```{r}
vars_out = filter(variables_excluded,
              studytype == 'Observational',
              outcome == 'actual') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

## Actual sample size for interventional studies

```{r}
vars_out = filter(variables_excluded,
              studytype == 'Interventional',
              outcome == 'actual') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

# ANZCTR

```{r, include=FALSE}
# get the results
load('results/ANZCTR_sample_size.RData') # from 2_elasticnet_model_samplesize_ANZCTR.R
```

## Target sample size for observational studies

```{r}
vars_out = filter(variables_excluded,
              study_type == 'Observational',
              outcome == 'target') %>%
  left_join(table_names_anzctr, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

## Target sample size for interventional studies

```{r}
vars_out = filter(variables_excluded,
              study_type == 'Interventional',
              outcome == 'target') %>%
  left_join(table_names_anzctr, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```


## Actual sample size for observational studies

```{r}
vars_out = filter(variables_excluded,
              study_type == 'Observational',
              outcome == 'actual') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

## Actual sample size for interventional studies

```{r}
vars_out = filter(variables_excluded,
              study_type == 'Interventional',
              outcome == 'actual') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```


# Sample size ratio

```{r, include=FALSE}
# get the results
load('results/sample_size_ratio.RData') # from 2_model_actual_target_ratio.R
```


## Observational studies

```{r}
vars_out = filter(variables_excluded,
              study_type == 'Observational',
              lambda == 'min') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

## Interventional studies

```{r}
vars_out = filter(variables_excluded,
              study_type == 'Interventional',
              lambda == 'min') %>%
  left_join(table_names_clintrials, by=c('variables'='term')) %>% # add nice label
  filter(!reference) %>% # exclude any stray reference categories
  select(group, label)
ftab = flextable(vars_out) %>%
	theme_box()%>%
	autofit()
ftab
```

